---
title: "Creating CSV"
author: "Team SIM"
date: "2023-04-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)


```

CSV files Drive folder : https://drive.google.com/drive/folders/10GxigC1SFeK1WMFMoA2KCFU3k7Jnnznd?usp=sharing 

if folder is not shared or not accessible please send a request access


The csv files are where downloaded from Capital Bikeshare website, where october 2010 to march 2020 have one set of variables and later ones have the other set. so two folder are created to read in two sets.

```{r}
# # Set the path to the folder containing the CSV files
# csv_folder <- "~/Desktop/Intro to Data Science/final project CaBi/cabi all csv/csv/1 set"
# 
# # Get a list of all CSV files in the folder
# csv_files <- list.files(path = csv_folder, pattern = "*.csv", full.names = TRUE)
# 
# # Initialize an empty list to store the data frames
# csv_list <- list()
# 
# # Loop through each CSV file and read it into a data frame
# for (i in 1:length(csv_files)) {
#   csv_list[[i]] <- read.csv(csv_files[i], header = TRUE)
# }
# 
# # Merge all the data frames into one
# merged_csv1 <- do.call(rbind, csv_list)

```


This reading CSVs from the folders took so much time to get it merged in one file, so its saved to local that can we used easily on other platforms
```{r}
# merged_csv1 <- merged_csv1[c(1,2,4,5,8,9)]
# write.csv(merged_csv1,file = "merged_csv1.csv",row.names = F)

```


```{r}
# Reads a CSV file and creates a data frame in R.
CaBi1 <- read.csv("merged_csv1.csv")
```
Result:The CSV file merged_csv1.csv is read and stored as a data frame named CaBi1 in the R environment.


```{r}
# Shows the first few rows of the data frame CaBi1.
head(CaBi1)
```
Result:Displays the first 6 rows of the data frame CaBi1, giving a glimpse of its contents.

```{r}
# Returns the number of rows in the data frame CaBi1.
nrow(CaBi1)
```
Result:Returns the number of rows in the data frame CaBi1. This number represents the total number of observations in the data set.


```{r}
# Removing the Start.station.number and Bike.number as they are not required for further process
CaBi1 <- CaBi1 %>% select(-Start.station.number)
CaBi1 <- CaBi1 %>% select(-Bike.number)
unique(CaBi1$Member.type)
```



```{r}
#for Hourly
# These commands prepare the date-time data in the `CaBi1` data frame for time-based analyses by converting the "Start.date" column to a POSIXct date-time object in Eastern Standard Time and rounding it to the nearest hour.
# CaBi1$Start.date <- strptime(CaBi1$Start.date,format="%Y-%m-%d %H:%M:%S",tz = "EST")

# CaBi1$Start.date <- format(round(CaBi1$Start.date, units="hours"), format="%Y-%m-%d %H:%M:%S")
```
Result:The "Start.date" column of the `CaBi1` data frame is rounded to the nearest hour and converted back to a character string in the format "%Y-%m-%d %H:%M:%S", which can be useful for aggregating the data by hour or for formatting the date-time data in a specific way.



```{r}
#For day wise start.date we removed the time
CaBi1$Start.date <- strptime(CaBi1$Start.date,format="%Y-%m-%d",tz = "EST")
```
Result: "Start.date" column of CaBi1 data frame is converted to a POSIXct date object.

```{r}
#Groups CaBi1 by "Start.date" and "Start.station", calculates number of bikes and average duration per group, keeps only distinct rows, and sorts by "Start.date" and "Start.station".
CaBi1 <- CaBi1 %>% 
  group_by(Start.date, Start.station) %>% 
  mutate(noofBikes = n(), 
         Duration = mean(Duration, na.rm = TRUE)) %>% 
  distinct(Start.date, Start.station, .keep_all = TRUE) %>% 
  arrange(Start.date,Start.station)
```
Result: CaBi1 is updated with "noofBikes" and "Duration" columns and sorted by "Start.date" and "Start.station".


```{r}
head(CaBi1)
```
Result: The first few rows of the CaBi1 data frame are displayed.



```{r}
# for hours
# write.csv(CaBi1,file = "CaBi1hour.csv",row.names = F)
```
Result: The CaBi1 data frame is written to a CSV file named "CaBi1hour.csv" without row names.



```{r}
# for days
write.csv(CaBi1,file = "CaBi1days.csv",row.names = F)
```
Result: The CaBi1 data frame is written to a CSV file named "CaBi1days.csv" without row names.



Doing the same with folder 2 which has the different set of variables
```{r}
# # Set the path to the folder containing the CSV files
# csv_folder <- "~/Desktop/Intro to Data Science/final project CaBi/cabi all csv/csv/2 set"
# 
# # Get a list of all CSV files in the folder
# csv_files <- list.files(path = csv_folder, pattern = "*.csv", full.names = TRUE)
# 
# # Initialize an empty list to store the data frames
# csv_list <- list()
# 
# # Loop through each CSV file and read it into a data frame
# for (i in 1:length(csv_files)) {
#   csv_list[[i]] <- read.csv(csv_files[i], header = TRUE)
# }
# 
# # Merge all the data frames into one
# merged_csv2 <- do.call(rbind, csv_list)
```
Result: The code reads multiple CSV files from a folder, stores them as individual data frames, and then merges all the data frames into a single data frame named "merged_csv2".

```{r}
# head(merged_csv2)
```
Result: The first few rows of the merged_csv2 data frame are displayed.


```{r}
# print(colSums(is.na(merged_csv2)))

```
Result: The sum of missing values for each column in the merged_csv2 data frame is printed.


```{r}
# merged_csv2 <- merged_csv2[c(3,4,5,7,13)]
```


```{r}
# head(merged_csv2)
# write.csv(merged_csv2,file = "merged_csv2.csv",row.names = F)
```
Result: The first few rows of the merged_csv2 data frame are displayed, and the merged_csv2 data frame is written to a CSV file named "merged_csv2.csv" without row names.

```{r}
CaBi2 <- read.csv("merged_csv2.csv")
```

```{r}
head(CaBi2)
```

```{r}
summary(CaBi2)
```



```{r}
CaBi2 <- CaBi2 %>% select(-end_station_name)

```
Result: The end_station_name column is removed from the CaBi2 data frame.


```{r}
CaBi2$Duration <- difftime(CaBi2$ended_at, CaBi2$started_at)

```
Result: A new column named Duration is added to the CaBi2 data frame containing the difference between the ended_at and started_at columns as a time duration.


```{r}
# This command removes the "ended_at" column from the CaBi2 data frame. It is likely that this column is no longer needed after calculating the duration of each trip in the previous step.
CaBi2 <- CaBi2 %>% select(-ended_at)
```


```{r}
#for Hourly
# CaBi2$started_at <- strptime(CaBi2$started_at,format="%Y-%m-%d %H:%M:%S",tz = "EST")
# 
# CaBi2$started_at <- format(round(CaBi2$started_at, units="hours"), format="%Y-%m-%d %H:%M:%S")
```
Result:The "started_at" column of the CaBi2 data frame is rounded to the nearest hour and converted back to a character string in the format "%Y-%m-%d %H:%M:%S". This can be useful for aggregating the data by hour or for formatting the date-time data in a specific way.




```{r}
#For day wise
CaBi2$started_at <- strptime(CaBi2$started_at,format="%Y-%m-%d",tz = "EST")

CaBi2$started_at <- format(CaBi2$started_at, "%Y-%m-%d")
```
Result:The "started_at" column of the CaBi2 data frame is converted from a character string to a POSIXct date-time object with time set to midnight (00:00:00) and timezone set to Eastern Standard Time, and then formatted to just the date ("%Y-%m-%d").


```{r}
# formatting the duration variable to double type

CaBi2$Duration <- as.double(CaBi2$Duration)
```


```{r}
CaBi2 <- CaBi2 %>% 
  group_by(started_at, start_station_name) %>% 
  mutate(noofBikes = n(),
         Duration = mean(Duration, na.rm = TRUE)) %>% 
  distinct(started_at, start_station_name, .keep_all = TRUE) %>% 
  arrange(started_at,start_station_name)
```
Result: This command converts the "Duration" column of the CaBi2 data frame from a time difference object to a numeric object. This can be useful for performing calculations and aggregations on the duration data. same as for the set 1 noofbikes is mutated with the group of started_at, start_station_name.


```{r}
# for hours
# write.csv(CaBi2,file = "CaBi2hour.csv",row.names = F)
```
Result:The data frame CaBi2 to a CSV file named "CaBi2hour.csv" without row names.


```{r}
# for days
write.csv(CaBi2,file = "CaBi2days.csv",row.names = F)
```
Result:The CaBi2 data frame to separate CSV file named CaBi2days.csv" with row names excluded. 
The CSV1 and CSV2 files are next used in the merging process.



