---
title: "csv"
author: "TEAM SIM"
date: "2023-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```


```{r}
CaBi1 <- read.csv("CaBi1days.csv")
CaBi2 <- read.csv("CaBi2days.csv")
```
Result: Reads two CSV files named "CaBi1days.csv" and "CaBi2days.csv" into two separate data frames named "CaBi1" and "CaBi2".


```{r}
head(CaBi1)
head(CaBi2)
```
Result: These two lines of code display the first six rows of data in the "CaBi1" and "CaBi2" data frames, respectively.



```{r}

colnames(CaBi1) <- c("duration", "started_at", "start_station_name","member_type","noofbikes")
colnames(CaBi2) <- c("started_at", "start_station_name","member_type","duration","noofbikes")

```
Result:Sets new column names for the "CaBi1" and "CaBi2" data frames, renaming the columns in the order "duration", "started_at", "start_station_name", "member_type", "noofbikes" for "CaBi1", and "started_at", "start_station_name", "member_type", "duration", "noofbikes" for "CaBi2".


```{r}
CaBi1 <- CaBi1[,c("started_at", "start_station_name","member_type","duration","noofbikes")]
```
Result: Rearranges the columns in the "CaBi1" data frame by selecting and reordering the columns in the order "started_at", "start_station_name", "member_type", "duration", "noofbikes".


```{r}
head(CaBi1)
head(CaBi2)
```

```{r}
CaBiday <- rbind(CaBi1, CaBi2)
head(CaBiday)
```
Result: Combines the "CaBi1" and "CaBi2" data frames into a new data frame called "CaBiday" with reordered columns. The `head()` function is used to display the first six rows of "CaBiday". 

```
CaBiday <- rbind(CaBi1, CaBi2)
head(CaBiday)
```
Result:Combines the "CaBi1" and "CaBi2" data frames into a new data frame called "CaBiday" with reordered columns. The head() function is used to display the first six rows of "CaBiday".



```{r}
write.csv(CaBiday,file = "CaBiday.csv",row.names = F)
```
Result:Writes the "CaBiday" data frame to a CSV file named "CaBiday.csv" in the current working directory, with row names excluded from the output.




```{r}
holidays <- read.csv("holiday.csv")
```
Result:Reads a CSV file named "holiday.csv" into a new data frame called "holidays".


```{r}
head(holidays)
```
```{r}
holidays <- holidays[c(1,3)]
holidays <- distinct(holidays, Date, .keep_all = TRUE)
nrow(holidays)
```
Result:Selects the first and third columns of the "holidays" data frame, removes duplicate rows based on the "Date" column, and counts the number of unique rows in the resulting data frame. 

```{r}
holidays$Date <- as.Date(paste0(holidays$Date, "-2022"), format = "%b %d-%Y")

# convert Date object to "dd-mm" format
holidays$Date <- format(holidays$Date, "%m-%d")
```
Result:Adds the year "2022" to the "Date" column of the "holidays" data frame and converts it to a date format. It uses the `as.Date()` function with the `paste0()` function to concatenate the year and the `format` argument to specify the input format.
Reformats the "Date" column of the "holidays" data frame from "YYYY-MM-DD" to "%m-%d" format using the format() function. The resulting date format only shows the month and day.



```{r}
holidays$Date <- strptime(holidays$Date,format="%m-%d",tz = "EST")
holidays$Date <- format(holidays$Date, "%m-%d")
```
Result:Converts the "Date" column of the "holidays" data frame from character to date-time format with EST timezone using strptime(). The second line reformats the "Date" column back to "%m-%d" format using format().



```{r}
#The `holidays` data frame is written to a CSV file named "holidays.csv" in the current working directory with row names excluded.
write.csv(holidays,file = "holidays.csv",row.names = F)
```

```{r}

# for weather 
# Set the path to the folder containing the CSV files
csv_folder <- "~/Desktop/Intro to Data Science/final project CaBi/cabi all csv/csv/weather "

# Get a list of all CSV files in the folder
csv_files <- list.files(path = csv_folder, pattern = "*.csv", full.names = TRUE)

# Initialize an empty list to store the data frames
csv_list <- list()

# Loop through each CSV file and read it into a data frame
for (i in 1:length(csv_files)) {
  csv_list[[i]] <- read.csv(csv_files[i], header = TRUE)
}

# Merge all the data frames into one
weather_csv <- do.call(rbind, csv_list)

```

```{r}
weather_csv <- weather_csv[c(2,5,8,9,10,18,25,30)]
write.csv(weather_csv,file = "weather_csv.csv",row.names = F)
```
Result:Selects specific columns from the `weather_csv` data frame and writes the resulting data frame to a CSV file named "weather_csv.csv".


```{r}
CaBiday <- read.csv("CaBiday.csv")
head(CaBiday)
weather <- read.csv("weather_csv.csv")
head(weather)
```
Result:Reading CaBiday.csv and weather_csv.csv files and displaying the top 6 rows of each file.


  
```{r}

weather$datetime <- strptime(weather$datetime,format="%Y-%m-%d",tz = "EST")

CaBiday$temp <- strptime(CaBiday$started_at,format="%Y-%m-%d",tz = "EST")
```
Result:Converts the started_at variable to Date type, which can be compared to the datetime variable in the weather dataset.




```{r}
#merge the 2 dataframes
CaBi <- merge(CaBiday,weather,by.x = "temp",by.y = "datetime")
```

```{r}
head(CaBi)
rm(weather)
rm(CaBiday)
rm(holidays)
```




* Weekend or Weekday
  * using existing tbl column add new column which will contain whether date is weekday or weekend
```{r}
CaBi$started_at <- strptime(CaBi$started_at,format="%Y-%m-%d")


CaBi$weekday <- ifelse(CaBi$temp$wday==0 | CaBi$temp$wday==6, "Weekend",  "Weekday")
```





* add holiday data from holiday list csv
```{r}
holidays <- read.csv("holidays.csv",header = T,stringsAsFactors = FALSE)
colnames(holidays) <- c("Date","Holiday")

CaBi$t <- format(as.Date(CaBi$temp), "%m-%d")

CaBi <- merge(CaBi,holidays,by.x = "t",by.y = "Date",all.x = TRUE)
```
Result:Reads "holidays.csv" file into a data frame named "holidays".Renames column names to "Date" and "Holiday".Merges "holidays" with "CaBi" data frame using "t" column from "CaBi" and "Date" column from "holidays". Save the result to "CaBi". Use all.x=TRUE to keep all the rows from the left data frame.

```{r}
head(holidays)
unique(holidays$Date)
```


```{r}
# for hour
# CaBi$workhour <- ifelse(CaBi$Start.date$hour>=9 & CaBi$Start.date$hour<=17,"Working Hour","Non-Working Hour")
```


* Season => Fall, winter, summer, spring 
```{r}
CaBi$season <- ifelse(CaBi$temp$mon %in% c(11,0,1), "Winter",
               ifelse(CaBi$temp$mon %in% c(2,3,4), "Spring",
               ifelse(CaBi$temp$mon %in% c(5,6,7), "Summer","Fall")))

```

```{r}
head(CaBi)
```


```{r}
CaBi$started_at <- strptime(CaBi$started_at,format="%Y-%m-%d",tz="EST")


CaBi$date <- format(CaBi$started_at,"%d")
CaBi$month <- format(CaBi$started_at, "%m")
CaBi$year <- format(CaBi$started_at, "%Y")
```


```{r}
head(CaBi)
```


```{r}
CaBi <- CaBi[, -c(1, 2)]

```

```{r}
colnames(CaBi) <- c("started_at","start_station_name","member_type","duration","noofbikes","temperature","feelsliketemp","dew","humidity","windspeed","uvindex","weather","weekday","holiday","season","date","month","year")

```




```{r}
write.csv(CaBi,file = "CaBi.csv",row.names = F)

```



