---
title: "csv"
author: "TEAM SIM"
date: "2023-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```


```{r}
CaBi1 <- read.csv("CaBi1days.csv")
CaBi2 <- read.csv("CaBi2days.csv")
```
Result: Reads two CSV files named "CaBi1days.csv" and "CaBi2days.csv" into two separate data frames named "CaBi1" and "CaBi2".


```{r}
head(CaBi1)
head(CaBi2)
```
Result: These two lines of code display the first six rows of data in the "CaBi1" and "CaBi2" data frames, respectively.



```{r}

colnames(CaBi1) <- c("duration", "started_at", "start_station_name","member_type","noofbikes")
colnames(CaBi2) <- c("started_at", "start_station_name","member_type","duration","noofbikes")

```
Result:Sets new column names for the "CaBi1" and "CaBi2" data frames, renaming the columns in the order "duration", "started_at", "start_station_name", "member_type", "noofbikes" for "CaBi1", and "started_at", "start_station_name", "member_type", "duration", "noofbikes" for "CaBi2".


```{r}
CaBi1 <- CaBi1[,c("started_at", "start_station_name","member_type","duration","noofbikes")]
```
Result: Rearranges the columns in the "CaBi1" data frame by selecting and reordering the columns in the order "started_at", "start_station_name", "member_type", "duration", "noofbikes".


```{r}
head(CaBi1)
head(CaBi2)
```

```{r}
CaBiday <- rbind(CaBi1, CaBi2)
head(CaBiday)
```
Result: Combines the "CaBi1" and "CaBi2" data frames into a new data frame called "CaBiday" with reordered columns. The `head()` function is used to display the first six rows of "CaBiday". 

```
CaBiday <- rbind(CaBi1, CaBi2)
head(CaBiday)
```
Result:Combines the "CaBi1" and "CaBi2" data frames into a new data frame called "CaBiday" with reordered columns. The head() function is used to display the first six rows of "CaBiday".



```{r}
write.csv(CaBiday,file = "CaBiday.csv",row.names = F)
```
Result:Writes the "CaBiday" data frame to a CSV file named "CaBiday.csv" in the current working directory, with row names excluded from the output.




```{r}
holidays <- read.csv("holiday.csv")
```


```{r}
head(holidays)
```
```{r}
holidays <- holidays[c(1,3)]
holidays <- distinct(holidays, Date, .keep_all = TRUE)
nrow(holidays)
```


```{r}
holidays$Date <- as.Date(paste0(holidays$Date, "-2022"), format = "%b %d-%Y")

# convert Date object to "dd-mm" format
holidays$Date <- format(holidays$Date, "%m-%d")
```



```{r}
holidays$Date <- strptime(holidays$Date,format="%m-%d",tz = "EST")
holidays$Date <- format(holidays$Date, "%m-%d")
```



```{r}
write.csv(holidays,file = "holidays.csv",row.names = F)
```







```{r}

# for weather 
# Set the path to the folder containing the CSV files
csv_folder <- "~/Desktop/Intro to Data Science/final project CaBi/cabi all csv/csv/weather "

# Get a list of all CSV files in the folder
csv_files <- list.files(path = csv_folder, pattern = "*.csv", full.names = TRUE)

# Initialize an empty list to store the data frames
csv_list <- list()

# Loop through each CSV file and read it into a data frame
for (i in 1:length(csv_files)) {
  csv_list[[i]] <- read.csv(csv_files[i], header = TRUE)
}

# Merge all the data frames into one
weather_csv <- do.call(rbind, csv_list)

```

```{r}
weather_csv <- weather_csv[c(2,5,8,9,10,18,25,30)]
write.csv(weather_csv,file = "weather_csv.csv",row.names = F)
```


```{r}
CaBiday <- read.csv("CaBiday.csv")
head(CaBiday)
weather <- read.csv("weather_csv.csv")
head(weather)
```


  
```{r}

weather$datetime <- strptime(weather$datetime,format="%Y-%m-%d",tz = "EST")

CaBiday$temp <- strptime(CaBiday$started_at,format="%Y-%m-%d",tz = "EST")
```




```{r}
#merge the 2 dataframes
CaBi <- merge(CaBiday,weather,by.x = "temp",by.y = "datetime")
```

```{r}
head(CaBi)
rm(weather)
rm(CaBiday)
rm(holidays)
```




* Weekend or Weekday
  * using existing tbl column add new column which will contain whether date is weekday or weekend
```{r}
CaBi$started_at <- strptime(CaBi$started_at,format="%Y-%m-%d")


CaBi$weekday <- ifelse(CaBi$temp$wday==0 | CaBi$temp$wday==6, "Weekend",  "Weekday")
```





* add holiday data from holiday list csv
```{r}
holidays <- read.csv("holidays.csv",header = T,stringsAsFactors = FALSE)
colnames(holidays) <- c("Date","Holiday")

CaBi$t <- format(as.Date(CaBi$temp), "%m-%d")

CaBi <- merge(CaBi,holidays,by.x = "t",by.y = "Date",all.x = TRUE)
```


```{r}
head(holidays)
unique(holidays$Date)
```


```{r}
# for hour
# CaBi$workhour <- ifelse(CaBi$Start.date$hour>=9 & CaBi$Start.date$hour<=17,"Working Hour","Non-Working Hour")
```


* Season => Fall, winter, summer, spring 
```{r}
CaBi$season <- ifelse(CaBi$temp$mon %in% c(11,0,1), "Winter",
               ifelse(CaBi$temp$mon %in% c(2,3,4), "Spring",
               ifelse(CaBi$temp$mon %in% c(5,6,7), "Summer","Fall")))

```

```{r}
head(CaBi)
```


```{r}
CaBi$started_at <- strptime(CaBi$started_at,format="%Y-%m-%d",tz="EST")


CaBi$date <- format(CaBi$started_at,"%d")
CaBi$month <- format(CaBi$started_at, "%m")
CaBi$year <- format(CaBi$started_at, "%Y")
```


```{r}
head(CaBi)
```


```{r}
CaBi <- CaBi[, -c(1, 2)]

```

```{r}
colnames(CaBi) <- c("started_at","start_station_name","member_type","duration","noofbikes","temperature","feelsliketemp","dew","humidity","windspeed","uvindex","weather","weekday","holiday","season","date","month","year")

```




```{r}
write.csv(CaBi,file = "CaBi.csv",row.names = F)

```



